Clouds and Radiation in a mock-Walker Circulation

### 2020-09-10 Silvers_WalkerCell_response2reviewers.tex

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

%% ------------------------------------------------------------------------ %%

\section{Note from LGS and TR}

\textcolor{blue}{All responses and comments by Levi Silvers will be in blue.  Below please find my responses to your particular considerations.  When referring to line numbers that were used in your revues I have done my best to use the same line numbers (those from the original submission) unless otherwise
noted.  Reviewer \#1 made additional comments in a pdf that was included with the review.  I have responded to those comments within that pdf.  Reviewer \#3 made 15 minor comments in the annotated pdf provided by \textit{JAMES}.  I have responded with my own annotations. }

@@

  \item Ln 338-343: Yes, but maybe clarify that you must compare similar domain sizes in Table 2. The variation in SF is also influenced by domain size. 
  \textcolor{red}{I agree in principle that the cleanest comparisons will be with similar domain sizes, but comparing similar domain sizes does not help to clarify the numbers in Table 2.  For example, both the largest and smallest SF are found on the large domain.  }

  \item Figure 6: I think there are too many lines on this plot. Maybe two separate plots would be best: one for P25 and one for P100.
  \textcolor{blue}{Done.  Splitting this into two panels was a good suggestion.}

  \item Ln 461: 'more' instead of 'better'
  \textcolor{blue}{Done}

  \item Figure 13: These plots are useful, however I still think it might be helpful to show a profile of the longwave heating rate vs. horizontal distance (like Fig. 10)
%
Also, the triangle markers for E25 may be obscuring some details. 
I wonder if just using a slightly different color for E25 and P25 would be better.
\textcolor{red}{Figure 13 has been redone as a 2x2 panel figure using a different color for E25 instead of the triangles, larger axis labels, and clearer panel titles.}

@@

  \item Ln 545-546: This is particularly true because the low clouds are found in relatively dry regions. Without low clouds, the dry atmosphere would emit less downwelling longwave radiation and roughly the same outgoing longwave radiation. In other words, the low clouds effectively increase the downwelling radiation in dry regions, which leads to more atmospheric cooling.   
  \textcolor{blue}{This is a good way of stating the underlying reasons.  I have made some modifications so the sentence now 
  reads, ``Low-level clouds, especially in dry regions, serve as a significant source of  radiative cooling for the atmosphere (Figure 
  13) because they increase the downwelling longwave radiation.''}

  \item Ln 556-557: A bit unclear. What decrease of low-level clouds are you referring to here? Are you referring to the fact that the CRM-like simulations have fewer low clouds?
  \textcolor{blue}{I have added, `when the LWCRE is off' to this sentence to make my meaning clear.  I am not referring to the fact that CRM-like simulations have fewer clouds, but to the fact that there are far less low-level clouds in the LWCRE-off simulations 
  especially for the GCM-like simulations. }

  \item L584-608: I think it is good to have a summary here, but I'm getting lost in all of the details. It would be helpful for the reader if you could summarize only the most important points (maybe 5-6 bullets max). For example, I think you could probably eliminate the last 2 points in the first set (i.e. about spatial gradients of humidity, etc and about the condensate changes).

You also do not summarize any of the domain size results here, however they are mentioned in your conclusions. So I'm left a bit confused about the overall importance of the domain size results... to remedy that I suggest 1) mentioning the key domain-size sensitivies here and/or 2)  deleting the discussion of the domain-size sensitivities in the conclusions. Just trying to help you better convey the main points to the reader.
  \textcolor{blue}{The last bullet point (Lines 592-594) of the first list has been removed.  I have also reorganized and combined the material that was 
  formerly in sections 6 and 7 into one section with the hope of more clearly distinguishing between the summary of our results and the discussion points.  Hopefully it is not as easy to get lost in the details now.}
  \textcolor{red}{I sympathize with you.  This paper has been hard to write because it has been tough to not get lost in all of the details myself.  The \textit{importance} of several of these results, such as the domain size results, is not clear to me, but they 
  seem both interesting and significant.}

@@

  \textcolor{blue}{In contrast to the P25 (lwcre-on) simulation (Figure 1, left; Figure 4d), the initial several months of both P25 
  (lwcre-off; Figure 5d) and E25 (lwcre-on; Figure 8b) show a precipitation maximum that is centered closely on the SST 
  maximum in the middle of the domain.  The steady state solutions from the equilibrated period of these simulations (Figure 1) are 
  also both more symmetric about the center of the domain.  This implies that the source of asymmetry could be either the lwcre, 
  or the parameterized convection.  However, both P25(lwcre-off) and E25(lwcre-on) eventually transition to a much more irregular evolution of precipitation (see Figure 5d after 600 days) in which the precipitation maximum can reside well off of the 
  precipitation maximum for several months at a time.  However, both of these simulations have a more symmetric time-averaged distribution than the P25 case.  I then made the inference that it is the interaction between the two that leads to an asymmetric distribution of the steady state. }

\textcolor{blue}{
This has been concluded mostly because the simulation is markedly more symmetrical when the LWCRE is off but the convection parameterization is still on, and the equilibrium state is more symmetrical when the LWCRE is on but the convection parameterization is off (Figure 1).  The time evolution of precipitation for the LWCRE-off, Conv-off (E25 lwcre off) also shows regular oscillations (large) about the SST maximum in a way that is distinct from either P25 lwcre-off or E25 lwcre-on.  Another piece of this (I was a bit hesitant to include too many Hovmoller plots, fearing Hovmoller fatigue.)  }
\textcolor{red}{include Hovmoller figure with four different 25km simulations?}

  \item Line 326-343: The focus in Table 2 is on the equilibrium state. Is a time evolution of subsidence fraction available? I'd be curious to see how SF changes from the early stages of the simulation, i.e. if it's becoming more or less aggregated over time. The somewhat inconsistent comparison between LWCRE-on and LWCRE-off experiments here is noteworthy, given that LWCRE is arguably the most important mechanism for maintenance of the aggregated state in non-rotating RCE studies.
\textcolor{blue}{Good question.  Once the experiments equilibrate the subsidence fraction (SF) is quite steady in time, despite rather large variability (on the order of 0.1-0.15).   Variability increases with grid-spacing to the point that for P100 experiments averages must be taken over at least 3-5 months to obtain the long term mean SF.
Figure \ref{fig:subsfracTS}  
shows the subsidence fraction for the control (LWCRE-on) experiments.  The LWCRE-off cases are also very steady in time with a similar degree of variability.  The mean 
differences of SF between the LWCRE on/off experiments are shown in an updated Table 2.  It should also be noted that SF in the case of the Walker 
Circulation should be expected to be relatively large, and more consistent across experiments (relative to pure RCE) because of the prescribed SST warm 
patch that will largely fix the location of deep convection to the middle of the domain and force an overturning circulation.  
The values used in Table 2 for the originally submitted manuscript were computed 
using the time averaged mid-tropospheric vertical velocity because I had neglected to save vertical velocity at a daily frequency.  However, I have since 
found that the daily vertical pressure velocity (omega) was saved on the 500 hPa pressure level.  I have recomputed the SF values using these omega values
and updated Table 2 (caption has been updated).  The SF values now consistently show that the LWCRE-off experiments have a lower SF, the values are more consistent across experiments, and the surprisingly large difference between the LWCRE off/on experiments for the P100 large domain experiments is reduced to a value of 0.1.  I have deleted 'with the exception of E25' from Line 332 of the submitted manuscript.}

\begin{figure}
  \centering
    \includegraphics[width=0.8\columnwidth]{walkerfigs/SubsFrac_TS.eps}  
  \caption{Subsidence Fraction over the first two years of simulation (E1 and E2 only extend for 6 months.).  Values computed from the pressure 
  velocity on the 500 mb pressure surface.}
  \label{fig:subsfracTS}
\end{figure}

@@

  \textcolor{blue}{It is due mostly to the changes in the latent heat flux (~5-10 W/m2) rather than the sensible heat flux (~1 W/m2).}

  \item Line 676-678: This can be a significant result that warrants future research. Has there been any other work into the sensitivity of the large-scale cloud scheme in GCMs besides Held et al. (2007)? If so, this would be useful to cite and elaborate on, either in this section or back in Section 4.

  \textcolor{blue}{I am not aware of other research that has looked into this.  This was one of the points that I found the moist interesting to come out of the analysis.  I am continuously surprised at how infrequently the large-scale vs. convective precipitation is talked about in studies that use GCMs.  I hope that future research does look into this further, whether my own or someone else’s.}

  \item Figure 13 caption: How are the subsidence regions defined? Specifically, is any sort of areal averaging conducted and does this vary between simulations of different domain size/resolution? Apologies if this is something that has already been addressed elsewhere in the manuscript.   
  \textcolor{blue}{Subsidence regions are defined as the quarter of the domain furthest from the SST 
  maximum.  In other words, the two sections, each an eighth of the domain width, which are at the outer edges of the domain.  I played around with this a bit, 
  especially because the asymmetry in some of the runs can complicate things.  But defining the subsidence region as only a quarter of the domain instead of, 
  say, half of the domain, minimizes the influence of asymmetry.  This method for defining the subsidence region is systematic across all experiments and while 
  there are some grid points with subsidence that are not included, the subsidence regions do not include grid-points with ascent.}

@@

This series of mock-Walker cell simulation is new, the manuscript is well-written, the results are consistent with past
literature, and the topic is of interest to the climate community as it fits well within the “climate model hierarchy” [3, 1],
all of which warrant eventual publication in the Journal of Advances in Modeling Earth Systems. I list a few comments
below and would be happy to help with another round of minor revisions if necessary.

\subsection{Major comments}
\subsubsection{Why limit the investigation of cloud radiative effect to LWCRE?}
While the climate dynamics and self-aggregation literature consistently find that the LWCRE is the most important
radiative forcing when it comes to convective aggregation (e.g. [6, 4, 2]), the consensus is less clear when it comes to
shortwave cloud radiative effect (SWCRE). Given the clean simulation setup and the insight obtained from the LWCRE-off
simulations, \textbf{would it be possible to repeat the experiments with SWCRE and/or the full cloud radiative
effect (CRE) turned off?}

I would recommend to:

\begin{itemize}
  \item Rerun at least one simulation with SWCRE-off or the full CRE-off, or
  \item Thoroughly justify why the authors decided to focus on LWCRE only rather than the full CRE. The paragraph
currently introducing the LWCRE-off experiments (L202) does not justify why the experiment is done in the first
place, even if the manuscript’s results confirm the importance of this experiment a posteriori.
\textcolor{blue}{Good question.  It is possible with this configuration to repeat experiments in which the SWCRE is turned off/on.  
We are using a very similar physics code as that used by \textit{Popp \& Silvers, 2017, J. of Clim.}.   However, we have chosen to limit ourselves in this case to the 
LWCRE experiments for several reasons, the foremost being practicality.  As you know, the current study varies LWCRE, resolution, domain size, and 
the convective parameterization.  As a result, organizing the current range of results is already a challenge and we chose not to 
expand the study further.  I have no doubt that the impact of the SWCRE on the coupling between the LS circulation and clouds will lead 
to more insight and I hope that it will be investigated in future studies.  Additional reasons for focusing our study on the LWCRE on/off experiments
are as follow.  First, this provides a clean overlap with CFMIP experiments which emphasize LWCRE over the SWCRE due to the complicating impacts of
SWCRE on the surface energy budget.  Second, there are numerous studies which have shown that the atmospheric shortwave heating is 
much smaller than the longwave heating (e.g. Perdergrass \& Hartmann, 2014; Popp \& Silvers, 2017 (their Figures 1 \& 4)).}
\textcolor{red}{This response should lean heavily on Popp and Silvers.  It should additionally emphasize the many directions which the paper already hints at due the large amount of experiments that are being compared, domain size differences, 
parameterization differences, resolution differences, and the LWCRE differences.}
\textcolor{blue}{draft of additions: Both the shortwave and longwave radiation influence the atmospheric radiative heating.  We have chosen to focus on 
the LWCRE because it is by far the dominant contributor to atmospheric radiative heating.
In order to present a relatively simple analysis of the cloud radiative effect we have chosen to explore LWCRE on/off experiments but not SWCRE on/off 
experiments.  }
\end{itemize}

### 2020-09-10 Silvers_WalkerCell_revision.tex
further modifications for the Walker Circulation paper in the files Silvers_WalkerCell_response2reviewers.tex Silvers_WalkerCell_revision.tex

\begin{table}
\begin{center}
\caption{Domain mean precipitation ($\overline{\rm{P}}$), outgoing longwave radiation 
($\overline{\rm{OLR}}$), precipitable water ($\overline{\rm{PW}}$), and subsidence fraction (SF)  
or fraction of domain that is subsiding at the 532 hPa level.
Values in parenthesis correspond to LWCRE-off experiments.}
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
    \\
 Name &   $\overline{\rm{P}} (\mathrm{mm \,d^{-1}})$ & $\overline{\rm{OLR}} (\rm{W \, m^{-2}})$ & $\overline{\rm{PW}} (\rm{mm})$ & SF   \\ \hline
  P100L   &  4.1 (3.5)   &  283.1 (286.9)  & 36.6 (31.3)  & 0.89 (0.61)      \\ 
    \\
  P100 &   3.9 (3.7)   &  283.2 (296.4)  & 28.0 (26.8) & 0.74 (0.73)           \\  
    \\
  P25L &   4.0 (3.8)  &  281.2 (290.7)   & 35.0 (32.9) & 0.78 (0.70)          \\  
    \\
  P25  &   3.8 (3.7)    & 282.9 (293.6)   & 27.4 (26.4) & 0.80 (0.74)          \\  
    \\
 E25  & 3.7 (3.5)    &   271.9 (286.8)   & 28.7 (27.3) & 0.72 (0.75)            \\  
    \\
 E2   &  3.1 (3.4)   &  266.2 (285.5)    & 27.0 (25.2) & 0.82 (0.75)           \\ 
    \\
 E1   &  3.3 (3.7)   &  269.3 (289.2)    & 27.3 (26.5) & 0.80 (0.80)         \\  \hline

@@

\begin{figure}
  \centering
      %\includegraphics[width=0.8\columnwidth]{walkerfigs/Precip_DMN_4models_1yr.eps}
      \includegraphics[width=0.8\columnwidth]{walkerfigs/DomMnPrecip.eps}
  \caption{Domain mean precipitation as a function of time.  All data have been smoothed twice with a 9-day running mean filter.  
  Solid lines are the control experiments, dashed lines show the LWCRE-off experiments.  All data is shown for the E1 and E2 experiments
  while for P25 and P100 only the first two out of five years is shown.  The dots at far right show the time mean values for the last 4 years (P25, P100) and last 5 months (E1, E2).  Filled (hollow) dots show values from control (LWCRE-off) experiments. }
    \label{fig:precip_dom_mn}
\end{figure}

One of the simplest measures of convective aggregation and the large-scale circulation is the subsidence fraction (SF), the fraction of the domain 
in which the mid-tropospheric air is subsiding \cite{Coppin_Bony_2015}.  
As convection becomes more organized, or aggregated, the SF will increase.  For an overturning circulation a contraction of the convective 
region should result in a larger subsidence fraction.  
This is precisely what we see in Table \ref{tab:experimentmeans}.  For each of our experiments with LWCRE-on the SF is larger than 
or equal to the case with LWCRE-off (with the exception of E25).   
While our mock-Walker circulation is distinct from radiative convective equilibrium (RCE) and the resulting spontaneous self-aggregation, our 
region of persistent deep convection clearly corresponds to a state of aggregation. 
Using a mock-Walker circulation allows one to study controlled convective aggregation rather then spontaneous convective self-aggregation. 
Previous studies have shown a dependence of aggregation 
on temperature \cite{Khairoutdinov2010, Wing2014, Cronin2017}, Table \ref{tab:experimentmeans} illustrates how much the aggregated 
state can vary among experiments with identical SST.  Having a prescribed SST warm 
patch ensures that the simulations will be `aggregated' to some degree.  
Given identical SSTs, the  range of different SFs provide a measure of variability that is driven entirely by the interactions between 
convection, radiation, and the large-scale circulation.

@@

\begin{figure}
  \centering       
    \includegraphics[width=0.96\columnwidth]{walkerfigs/lowlevel_New.eps}
    \caption{Low-level structure and domain mean wind shear for simulations with explicit convection.  
    The surface enthalpy flux (a) is the latent plus sensible heat flux.  Panels b and c show 
    the equivalent potential temperature (b) and zonal wind (c) at the lowest model level in the atmosphere.  Panel 
    d shows the domain mean zonal wind throughout the depth of the domain.}
    \label{fig:enthalpy}
\end{figure}

@@

\begin{figure}
  \centering
      \includegraphics[width=0.96\columnwidth]{walkerfigs/CloudFractdtRad_lwcre.eps}
  \caption{Cloud fraction (CF) and temperature tendency due to longwave radiation (tdtlw).  Profiles were computed in the 
  subsidence regions and are shown for the control (LWCRE-on) and LWCRE-off experiments.}
  \label{fig:cf_tdtlw}
\end{figure}

@@

\begin{itemize}
  \item An increase of $\overline{\rm{P}}$  for the P100,P25 and E25 experiments.
  \item A decrease of $\overline{\rm{P}}$  for the E1 and E2 experiments.
  \item The horizontally oriented low-level circulations strengthen, and the regions of high precipitation are more concentrated  
  in geographic space (Figures \ref{fig:rh_psi_P25vsE25}, \ref{fig:rh_psi_P25E2E1} and \ref{fig:rh_psi_P25E2E1_lwoff}).  
  \item The surface enthalpy flux decreases by tens of $\rm{W m^{-2}}$ and is much more variable among the models over the warm patch (Figure \ref{fig:enthalpy} a).
  \item The spatial gradients of $\theta_e$ (Figure \ref{fig:enthalpy} b), the specific humidity, temperature, and virtual temperature 
  (not shown) increases on the lowest atmospheric level.  
  \item When convection is parameterized the total domain mean condensate increases at all heights 
  (Figure \ref{fig:TotCond_P25P100}), in contrast to when the convection is explicit and 
  the condensate aloft decreases (Figure \ref{fig:TotCond}).    
\end{itemize}

>>>>

\begin{table}
\begin{center}
\caption{Domain mean precipitation ($\overline{\rm{P}}$), outgoing longwave radiation 
($\overline{\rm{OLR}}$), precipitable water ($\overline{\rm{PW}}$), and subsidence fraction (SF)  
or fraction of domain that is subsiding at the \change[LGS]{532}{500} hPa level.
Values in parenthesis correspond to LWCRE-off experiments.}
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
    \\
 Name &   $\overline{\rm{P}} (\mathrm{mm \,d^{-1}})$ & $\overline{\rm{OLR}} (\rm{W \, m^{-2}})$ & $\overline{\rm{PW}} (\rm{mm})$ & SF   \\ \hline
  P100L   &  4.1 (3.5)   &  283.1 (286.9)  & 36.6 (31.3)  & 0.82 (0.72)      \\ 
    \\
  P100 &   3.9 (3.7)   &  283.2 (296.4)  & 28.0 (26.8) & 0.74 (0.74)           \\  
    \\
  P25L &   4.0 (3.8)  &  281.2 (290.7)   & 35.0 (32.9) & 0.81 (0.78)          \\  
    \\
  P25  &   3.8 (3.7)    & 282.9 (293.6)   & 27.4 (26.4) & 0.82 (0.80)          \\  
    \\
 E25  & 3.7 (3.5)    &   271.9 (286.8)   & 28.7 (27.3) & 0.82 (0.81)            \\  
    \\
 E2   &  3.1 (3.4)   &  266.2 (285.5)    & 27.0 (25.2) & 0.86 (0.81)           \\ 
    \\
 E1   &  3.3 (3.7)   &  269.3 (289.2)    & 27.3 (26.5) & 0.82 (0.79)         \\  \hline

@@

\begin{figure}
  \centering
      %\includegraphics[width=0.8\columnwidth]{walkerfigs/Precip_DMN_4models_1yr.eps}
      \includegraphics[width=0.8\columnwidth]{walkerfigs/DomMnPrecip.eps}
  \caption{Domain mean precipitation as a function of time.
  Solid lines are the control experiments, dashed lines show the LWCRE-off experiments.  All data is shown for the E1 and E2 experiments
  while for P25 and P100 only the first two out of five years is shown.  The dots (control) and diamonds (LWCRE-off) at far right show the time 
  mean values of precipitation for the 
  last 4 years (P25, P100) and last 5 months (E1, E2).  The mean values are also given in the legend. 
  All data have been smoothed twice with a 9-day running mean filter.  }
    \label{fig:precip_dom_mn}
\end{figure}

One of the simplest measures of convective aggregation and the large-scale circulation is the subsidence fraction (SF), the fraction of the domain 
in which the mid-tropospheric air is subsiding \cite{Coppin_Bony_2015}.  
As convection becomes more organized, or aggregated, the SF will increase.  For an overturning circulation a contraction of the convective 
region should result in a larger subsidence fraction.  
This is precisely what we see in Table \ref{tab:experimentmeans}.  For each of our experiments with LWCRE-on the SF is larger than 
or equal to the case with LWCRE-off \remove[LGS]{(with the exception of E25)}.   
While our mock-Walker circulation is distinct from radiative convective equilibrium (RCE) and the resulting spontaneous self-aggregation, our 
region of persistent deep convection clearly corresponds to a state of aggregation. 
Using a mock-Walker circulation allows one to study controlled convective aggregation rather then spontaneous convective self-aggregation. 
Previous studies have shown a dependence of aggregation 
on temperature \cite{Khairoutdinov2010, Wing2014, Cronin2017}, Table \ref{tab:experimentmeans} illustrates how much the aggregated 
state can vary among experiments with identical SST.  Having a prescribed SST warm 
patch ensures that the simulations will be `aggregated' to some degree.  
Given identical SSTs, the  range of different SFs provide a measure of variability that is driven entirely by the interactions between 
convection, radiation, and the large-scale circulation.

@@

\begin{figure}
  \centering       
    \includegraphics[width=0.96\columnwidth]{walkerfigs/lowlevel_New_b.eps}
    \caption{Low-level structure and domain mean wind shear for simulations with explicit convection.  
    The surface enthalpy flux (a) is the latent plus sensible heat flux.  Also shown are the  
    equivalent potential temperature (b) and zonal wind (c) at the lowest model level in the atmosphere.  Panel 
    d shows the domain mean zonal wind throughout the depth of the domain.}
    \label{fig:enthalpy}
\end{figure}

@@

\begin{figure}
  \centering
      %\includegraphics[width=0.96\columnwidth]{walkerfigs/CloudFractdtRad_lwcre.eps}
      \includegraphics[width=0.96\columnwidth]{walkerfigs/CF_tdtlw_4pan_fig.eps}
  \caption{Cloud fraction and temperature tendency due to longwave radiation in subsidence regions.  
  Top panels (a. and b.) show results from the control experiments while lower panels (c. and d.) show the
  analogous LWCRE-off experiments.  Note the change of axis in the lower panels.}
  \label{fig:cf_tdtlw}
\end{figure}

@@

\begin{itemize}
  \item {An increase of $\overline{\rm{P}}$  for the P100,P25 and E25 experiments.}
  \item {A decrease of $\overline{\rm{P}}$  for the E1 and E2 experiments.}
  \item {The horizontally oriented low-level circulations strengthen, and the regions of high precipitation are more concentrated  
  in geographic space (Figures \ref{fig:rh_psi_P25vsE25}, \ref{fig:rh_psi_P25E2E1} and \ref{fig:rh_psi_P25E2E1_lwoff}).  }
  \item {The surface enthalpy flux decreases by tens of $\rm{W m^{-2}}$ and is much more variable among the models over the warm patch (Figure \ref{fig:enthalpy} a).}
  \item {The spatial gradients of $\theta_e$ (Figure \ref{fig:enthalpy} b), the specific humidity, temperature, and virtual temperature 
  (not shown) increases on the lowest atmospheric level.  }
 \item {\remove[LGS]{When convection is parameterized the total domain mean condensate increases at all heights 
  (Figure 6), in contrast to when the convection is explicit and 
  the condensate aloft decreases (Figure 11.}    }
\end{itemize}
